---
title: "nox_project"
output: pdf_document
date: "2024-04-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, include=FALSE}
#loading packages
library(pacman)
p_load(dplyr, tidyverse, haven, ggplot2, magrittr, broom, plm, did, lmtest, miceadds)

#loading data 
nbp_data = read_dta('nbp.dta')
nbp_avg = nbp_data %>%
  group_by(year, summer, nbp) %>%
  summarise(nox_emit = sum(nox_emit, na.rm = T))
```

``` {r, include = FALSE}
#pre-treatment
nbp_pre_data = nbp_data %>%
  filter(post == 0)

#nbp avg treatment = 1
nbp_avg1 = nbp_avg %>%
  filter(nbp == 1)
#nbp avg treatment = 0 
nbp_avg2 = nbp_avg %>%
  filter(nbp == 0)
#nbp raw treatment = 1
nbp_data1 = nbp_data %>%
  filter(nbp == 1)
#nbp raw treatment = 0
nbp_data2 = nbp_data %>%
  filter(nbp == 0)
```   
Variable “nbp” indicates whether or not the state was regulated by the NOx Budget Trading Program. “nbp” = 1 denotes that the state is regulated by the NBP program.

Variable “summer” indicates whether or not the observed time period was during summer months or during the off season. “summer” = 1 designates the observation occurred through summer months. 

The variable “post” = 1 indicates that the time period of observation is post-treatment. For the purposes of this study, “post” = 1 denotes 2003 and following years. 

```{r, echo=FALSE}
#making graphs to denote NOx in treated states
#Calculate mean NOx emissions for summer and non-summer months
mean_nox = aggregate(nox_emit ~ year + summer, data = nbp_data, FUN = mean)

#Plot mean NOx emissions over time
#both nbp = 1 & nbp = 0 states
plot1 = ggplot(nbp_avg, aes(x = year, y = nox_emit/1000, color = factor(summer))) +
  geom_line() +
  facet_wrap(~ nbp) +
  labs(x = "Year", y = "Mean NOx Emissions") +
  scale_x_continuous(breaks = seq(1997, 2008, by = 1)) + 
  scale_color_manual(values = c("blue", "red"), labels = c("Non-Summer", "Summer")) +
  theme_classic()
plot1
#nbp = 1
plot2 = ggplot(nbp_avg1, aes(x = year, y = nox_emit/1000, color = factor(summer))) +
  geom_line() +
  labs(x = "Year", y = "Mean NOx Emissions") +
  scale_x_continuous(breaks = seq(1997, 2008, by = 1)) + 
  scale_color_manual(values = c("blue", "red"), labels = c("Non-Summer", "Summer")) +
  theme_classic()
plot2
#nbp = 0
plot3 = ggplot(nbp_avg2, aes(x = year, y = nox_emit/1000, color = factor(summer))) +
  geom_line() +
  labs(x = "Year", y = "Mean NOx Emissions") +
  scale_x_continuous(breaks = seq(1997, 2008, by = 1)) + 
  scale_color_manual(values = c("blue", "red"), labels = c("Non-Summer", "Summer")) +
  theme_classic()
plot3
```
The Parallel Trends assumption states that in the absence of the treatment or intervention being studied, the average outcomes of the treatment and control groups would follow parallel trends over time. In other words, before the treatment is introduced, the trends in outcomes for the treatment group and the control group should be similar or “parallel” to each other.

Based on visual observation, the parallel trends assumption is likely to hold given the corresponding graph of states not participating in the NBP program. This graph shows that the trend between summer and winter months remains approximately similar in states without “treatment” from participating in the NBP program.

Moreover, based on the regression design with output variable as ‘nox_emit’ and independent variables ‘summer’, ‘year’, and ‘summer*year’, we find the data upholds the parallel trends assumption. This is demonstrated in the summer*year variable being statistically insignificant. Here, the statistical insignificance represents no discernable change between summer and winter months during the pre-treatment period, demonstrating the parallel trends assumption holds.

The goal of panel 0 is to act as a placebo test. The non-NBP states function as a control to determine whether or not the NBP policy is effective. By creating a graph that compares emissions in non-NBP states in corresponding time periods, the research demonstrates that the NBP policy had an impact on NOx emissions, relative to both prior summertime emissions, and the ‘control group’ (non-NBP states).

```{r, echo=FALSE}
#nbp states first

#calculate statistical test of parallel trends
#Define the DiD model formula
nbp_nox_did_regr1 = lm(nox_emit ~ summer + year + year:summer, 
                       cluster = "state_season_cluster",
                       data = nbp_pre_data)
summary(nbp_nox_did_regr1)

#calculate difference-in-difference regression for panel A
#non-clustered regression 
nbp_nox_did_regr2 = lm(nox_emit ~ summer + post + summer*post, 
                   data = nbp_data1)
summary(nbp_nox_did_regr2)

#creating item to cluster 
#questionable rationale to create this cluster --> more reasonable cluster would be at pure state level
nbp_data1$state_season_cluster = paste(nbp_data1$fips_state, nbp_data1$summer, sep = "_")

#clustered regression
nbp_nox_did_regr_cluster = lm.cluster(nox_emit ~ summer + post + summer*post,
                                   cluster = 'state_season_cluster',
                                   data = nbp_data1)
nox_cluster_df = data.frame(summary(nbp_nox_did_regr_cluster))

```
The Difference in Difference Regression is defined as ‘nox_emit’ regressed on ‘summertime’ , ‘post’ , and the interaction term ‘summer*post’ . This interaction term is defined as the effect of the NBP program post-2003, during the summer time. Our coefficient (-0.373169) on ‘summer*post’ indicates that the NBP program reduces the amount of NOx emissions in the observed periods post-2003 during the summer time. 


```{r, echo=FALSE}
#non-nbp states DID model
#create item to cluster
nbp_data2$state_season_cluster = paste(nbp_data2$fips_state, nbp_data2$summer, sep = "_")

#clustered regression
non_nbp_nox_did_regr_cluster = lm.cluster(nox_emit ~ summer + post + summer*post,
                                      cluster = 'state_season_cluster',
                                      data = nbp_data2)
non_nox_cluster_df = data.frame(summary(non_nbp_nox_did_regr_cluster))
```

The Difference in Difference Regression is again defined as ‘nox_emit’ regressed on ‘summer’ , ‘post’ , and the interaction term ‘summer*post’ . The difference in the regression lies in the data used. Previously, we used data with NBP-participating states. For this regression, we used non-NBP participating states. This interaction term is similarly defined as the effect of the NBP program post-2003 during the summer time. Our coefficient (-0.04215831) on ‘summer*post’ demonstrates a small effect on states in this dataset. This is expected given the non-NBP status of states in this regression.

```{r, include=FALSE}
#triple diff regression
#create item to cluster
nbp_data$state_season_cluster = paste(nbp_data$fips_state, nbp_data$summer, sep = "_")

#triple-diff regression
full_nbp_nox_did_regr_cluster = lm.cluster(nox_emit ~ summer + post + nbp + summer*post + summer*nbp + post*nbp + summer*post*nbp,
                                           cluster = 'state_season_cluster',
                                           data = nbp_data)
full_nox_cluster_df = data.frame(summary(full_nbp_nox_did_regr_cluster))

#estimate total NOx reduction over time = TE from triple-diff * time duration (2003 - 2007)
```

```{r}
full_nox_cluster_df
```

The Triple-Difference regression is defined as ‘nox_emit’ regressed on the variables ‘summer’ , ‘post’ , ‘nbp’ , and the interaction terms ‘summer*post’ , ‘summer*nbp’ , ‘post*nbp’ , and ‘nbp*summer*post’ . The interaction term ‘nbp*summer*post’ is similar to prior observed interaction terms, differing by accounting for the effect of NBP in summer post-treatment. The coefficient is also statistically different from 0 (-0.33101040, P-Value: 2.591020 e-3), meaning that for observations treated (nbp) post-treatment (post) during summertime (summer) there is a decline in NOx emissions. 

```{r, include=FALSE}
#re-estimate model as 2-way fixed effects model
out = lm(data = nbp_data1,
            nox_emit ~ summer*post + factor(fips_state) + factor(year),
                               cluster = ~fips_state)

summary(out)

nbp_data1$residuals = nbp_data1$nox_emit - predict(out, newdata = nbp_data1)
nbp_data1$prediction = predict(out, newdata = nbp_data1)

#new graph demonstrating fixed-effects model based on state, year, and treatment
plot4 = ggplot(data = nbp_data1,
       aes(x = year,
           y = prediction,
           color = factor(fips_state))) +
geom_line () +
facet_wrap(~ summer) +
theme(legend.position = "none")
plot4
```

```{r}
summary(out)
plot4
```


The diff-in-diff-in-diff model re-estimated as a 2-Way Fixed Effect model gives us the same outcome coefficients on the summer*post variable. We get the same outcome because we are accounting for the idiosyncratic differences between states and years, similar to the difference in difference models which compare the treatment across groups and time. The 2-Way Fixed Effects model also captures this effect through the incorporation of the factor variables into the model. This is also demonstrated in ‘plot4’ as the treatment is equal across time in all groups. 
If there were differences in treatment implementation or at different points in time, the 2-Way Fixed Effects model would not accurately capture the effect of treatment across groups or time.

